Collocated Snapshot Backup with VM-Centric Deduplication in the Cloud

### 2013-09-25 snapshot_system/fast14_deduplication.tex

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{Jin2009},
we conduct additional local similary guided deduplication on a snapshot by comparing
chunk fingerprints of a dirty segment 
with those in  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their content signature is the same.
This segment content signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). Note that this definition of
content similarity is an approximation~\cite{resemblance97}.  When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot recipe.  Then recipes of the similar segments are loaded to memory,
which contain chunk fingerpints to be compared.
To control the time cost of search, we set a limit on the the number of  similar segments recipes to be fetched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For a 2MB segment, its segment recipe is roughly 19KB which contains about 500 chunk fingerprints and other chunk metadata,
by limiting at most 10 similar segments to search, the amount of memory for maintaining those similar segment recipes is small.
As part of our local duplicate search we also compare the current segment
against the parent segment at the same offset.

%\item 
\textbf{Global deduplication with popular chunks and replication support.}
This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
Our key observation is that the local deduplication has removed most of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the top $k$
most popular items. This dataset is called the \textbf{PDS} (popular data set). 
We define chunk popularity as the number of unique copies of the chunk in the data-store,
that is the number of copies of the chunk after local deduplication.
This number can be computed periodically (e.g., on a weekly basis).
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk fingerprints (called \textbf{PDS index}) in a distributed shared memory.  
previous work~\cite{WeiZhangIEEE} has studied a solution taking advantage of data popularity. This
paper improves this with a more comprehensive scheme with similarity-guided search and fault isolation.

@@

\begin{table}[htbp]
\centering
\begin{tabular}{|p{1.25cm}|p{5.5cm}|}
\hline
$k$ &  the number of top most popular chunks selected for deduplication\\ 
\hline
$c$ &  the total amount of data chunks in a cluster of VMs\\ 
\hline
$c_u$ &  the total amount of unique fingerprints after perfect  deduplication\\
\hline
$f_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$\delta$ &  the percentage of duplicates detected in local deduplication\\
\hline
$\sigma$ & =$\frac{k}{c_u}$ which is  the percentage of unique data  belong to  PDS\\
\hline
$p$ & the number of machines in the cluster\\
% \hline
% $V$ & the average number of VMs hosted on each machine\\
\hline
$E$ & the deduplication efficiency of VC approach\\
%\hline
%$D$ & the amount of unique data on each machine\\
%\hline
%$C$ & the average data chunk size. Our setting is  4K.\\
%\hline
%$s$ & the average size of file system blocks in the distributed file system. The default is  64MB.\\
%\hline
%$m$ & memory size on each node used by VC\\ 
%\hline
%$E$ & the size of an popular data index entry\\
\hline
$N_1$ & the average number  of non-PDS file system blocks  in a VM for VC\\
\hline
$N_2$ & the average number  of PDS file system blocks  in a VM for VC\\
\hline
$N_o$ & the average number  of file system blocks  in a VM for VO\\
\hline
A(r) & the availability of a file system block with replication degree $r$\\
\hline
\end{tabular}
\caption{Modeling  parameters}
\label{tab:symbol}
\end{table}

@@

Deduplication efficiency of the VC approach using top $k$ popular chunks
is the percentage of duplicates that can be detected:  
\begin{equation}
\label{eq:dedupeff}
%\begin{split}
%e_k &= 
E=\frac{ c\delta + c(1-\delta) \sigma^{1-\alpha}}
{c  - c_u }.\\
%\end{split}
\end{equation}

@@

\begin{figure}[htbp]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={PDS Coverage},
            width=\linewidth,
            height=0.6\linewidth,
            cycle multi list={
                mline\nextlist
                [3 of]mmark*\nextlist
            },
            %cycle list name=mcolor,
            xlabel={Total num. chunks stored (in billions)},
            ylabel={PDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            mark options=solid,
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.30)},
            anchor=north},
            ]
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A4] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T4] {figures/cds_coverage.txt};
            \legend{Measured ($\sigma=1\%$),Measured ($\sigma=2\%$),Measured ($\sigma=4\%$),Predicted ($\sigma=1\%$),Predicted ($\sigma=2\%$),Predicted ($\sigma=4\%$)};
            \end{axis}
    \end{tikzpicture}
  \caption{Predicted vs. actual PDS coverage as data size increases.}
  \label{fig:cds-coverage}
\end{figure}
Ratio $\sigma^{1-\alpha}$ represents the percentage of the remaining
chunks detected as duplicates in global deduplication due to PDS.
We call this PDS coverage.
Figure~\ref{fig:cds-coverage} shows predicted PDS coverage using $\sigma^{1-\alpha}$ when $\alpha$ is fixed at
0.65 and measured PDS coverage in our test dataset.
$\sigma=2\%$ represents memory usage of approximately 100MB memory per machine for the PDS.
While the predicted value remains flat, the  real coverage actually increases as the data size
increases, which indicates that $\alpha$ increases with the data size. 
Therefore in practice, the coverage of the PDS increases as more VMs are involved.

\subsection{ Impact on Storage Space and Fault Isolation}

@@

The replication degree of the backup storage 
is $r$ for regular file blocks and $r=3$ is a typical setting in distributed
file systems~\cite{googlefs03,hdfs10}.
In the VC approach, a special replication degree $r_c$ is used for PDS data where $r_c>r$. 
The storage cost for VO with full deduplication is $c_u *r$ and for VC, it is
$ k*r_c  + (c-E(c-c_u))*r$. Thus The storage cost ratio of $VC$ over $VO$ is 
\[
\sigma \frac{r_c}{r} + \frac{c-e(c-c_u)}{c_u}.
\]
Since $\sigma$ is small (it is 2\% in our experiments),  
term $\sigma \frac{r_c}{r}$ is small in the above expression.  
Thus the impact on storage increase is very small even when we choose a large $r_c/r$ ratio. 
For example, $r_c/r=2$ or 3.

Next we  assess  the impact of losing $d$ machines 
to the the VC and VO approaches.  
A large $r_c/r$ ratio can have a positive impact on full availability of VM snapshot blocks.
We use file system blocks rather than a deduplication
data chunk as our unit of failure because the DFS keeps
filesystem blocks as its base unit of storage.
To compute the full availability of all snapshots of a VM, we derive
the probability of losing a snapshot file block of a VM by
estimating the number of file system blocks per VM in each approach.
As illustrated in Figure~\ref{fig:share},
we build a bipartite graph representing the association from unique file system blocks
to their corresponding VMs in each approach. An association edge is  drawn  from a file block  to a VM 
if this block is used by the VM.

For VC, each VM has an average number of $N_1$ non-PDS file system blocks
and  has  an average of  $N_2$ PDS file system blocks. 
Each non-PDS file system block is associated with one VM
 and  we denote that PDS file system blocks are
shared by an average of $V_c$ VMs. Let $s$ be the average number of chunks per file block, 
$V$ be the average number of VMs hosted on each machine. Then, 
\[
V *p*N_1 *s  \approx c-E(c-c_u)\; \mbox{ and } \; 
V *p*N_2 *s  \approx c_u \sigma V_c.
\]

For VO, each VM has an average of $N_o$ file system blocks
and let $V_o$ be the average number of VMs shared by each file system block.
\[
V *p*N_o *s  = c_u  V_o.
\]

Since each file block (with default size $64MB$) contains many chunks (on average 4KB),
each file block contains the hot low-level chunks shared by many VMs, and it also contains
rare chunks which are not shared.  From the above equations:
\[
\frac{N_1}{N_o} \approx  \frac{c-E(c-c_u)}{c_uV_o}.
\]

If we achieve very high deduplication efficiecny, then $N_1$ would be much smaller than $N_o$. 
When there is a failure in file blocks with replication degree $r$
and there is no failure for PDS data with more replicas,   a VM in
the VC approach has a much lower chance to lose a snapshot than VO. 
%Figure~\ref{fig:fsb-links} shows the average number of VMs sharing each file block from
%our test dataset.
%For VC, we only show
%the sharing degree of  PDS-oriented file blocs  because non-PDS blocks always belong to
%one VM.
%Figure~\ref{fig:fsb-links} shows the average number of VMs sharing each file block from our test data.
Figure~\ref{fig:vm-links} shows the average number of file system blocks for each VM in VC and in VO. 
From data, we have observed that $N_1 +N_2 < N_o$ in practice. 
This is likely because the PDS FSBs tightly pack data used by many VMs, 
which decreases the overall number of FSBs required to backup a VM.
Note that  if  the backup for multiple VMs were conducted concurrently, there could be many more
VMs sharing each file block on average in VO.

@@

\item
When $r_c \leq d$, both non-PDS and PDS file blocks in VC can have a loss.
The full snapshot availability of  a VM in the VC approach is
\begin{equation}
\label{eq:VC2}
% (1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
% *
% (1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2}.
A(r)^{N_1} * A(r_c)^{N_2}
\end{equation}
Table~\ref{tab:fsb-availability} lists the $A(r)$ values with
%that the availability of an individual file system block
with different replication degrees. This demonstrates inequality 
% $1-\frac{ \binom{d}{r}} { \binom{p}{r} } < 1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} }$
$A(r) < A(r_c)$.  Since $N_1$ is much smaller than $N_o$ as discussed previously, 
these two conditions support that VC has a  higher availability. 
%comparing  Formula~\ref{eq:VO} and~\ref{eq:VC}. 
\end{itemize}

>>>>

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{Jin2009},
we conduct additional local similarity guided deduplication on a snapshot by comparing
chunk fingerprints of a dirty segment 
with those in  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their content signature is the same.
This segment content signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot metadata (called recipe). Note that this definition of
content similarity is an approximation~\cite{resemblance97}.  When processing a dirty segment,
its  similar segments can be found easily from the
parent snapshot recipe.  Then recipes of the similar segments are loaded to memory,
which contain chunk fingerprints to be compared.
To control the time cost of search, we set a limit on the the number of  similar segments recipes to be fetched. 
% Then,
%given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
%with those in similar segments.  
For a 2MB segment, its segment recipe is roughly 19KB which contains about 500 chunk fingerprints and other chunk metadata,
by limiting at most 10 similar segments to search, the amount of memory for maintaining those similar segment recipes is small.
As part of our local duplicate search we also compare the current segment
against the parent segment at the same offset.

%\item 
\textbf{Global deduplication with popular chunks and replication support.}
This step accomplishes the canonical global fingerprint lookup using a popular fingerprint index.
Our key observation is that the local deduplication has removed most of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the top $k$
most popular items. This dataset is called the \textbf{PDS} (popular data set). 
We define chunk popularity as the number of unique copies of the chunk in the data-store,
that is the number of copies of the chunk after local deduplication.
This number can be computed periodically (e.g., on a weekly basis).
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk fingerprints (called \textbf{PDS index}) in a distributed shared memory.  
Compared to a preliminary solution which  uses data popularity~\cite{WeiZhangIEEE}, 
this paper provides  a more comprehensive scheme with  improved deduplication efficiency and fault tolerance, and
analytic design guidance.

@@

\begin{table}[htbp]
\centering
\begin{tabular}{|p{1.25cm}|p{5.5cm}|}
\hline
$k$ &  the number of top most popular chunks selected for deduplication\\ 
\hline
$c$ &  the total amount of data chunks in a cluster of VMs\\ 
\hline
$c_u$ &  the total amount of unique fingerprints after perfect  deduplication\\
\hline
$f_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$\delta$ &  the percentage of duplicates detected in local deduplication\\
\hline
$\sigma$ & =$\frac{k}{c_u}$ which is  the percentage of unique data  belonging to  PDS\\
\hline
$p$ & the number of machines in the cluster\\
% \hline
% $V$ & the average number of VMs hosted on each machine\\
\hline
$E_c, E_o$ & deduplication efficiency of VC and VO \\
%\hline
%$D$ & the amount of unique data on each machine\\
%\hline
%$C$ & the average data chunk size. Our setting is  4K.\\
%\hline
%$s$ & the average size of file system blocks in the distributed file system. The default is  64MB.\\
%\hline
%$m$ & memory size on each node used by VC\\ 
%\hline
%$E$ & the size of an popular data index entry\\
\hline
$N_1$ & the average number  of non-PDS file system blocks  in a VM for VC\\
\hline
$N_2$ & the average number  of PDS file system blocks  in a VM for VC\\
\hline
$N_o$ & the average number  of file system blocks  in a VM for VO\\
\hline
A(r) & the availability of a file system block with replication degree $r$\\
\hline
\end{tabular}
\caption{Modeling  parameters}
\label{tab:symbol}
\end{table}

@@

Deduplication efficiency of the VC approach using top $k$ popular chunks
is the percentage of duplicates that can be detected:  
\begin{equation}
\label{eq:dedupeff}
%\begin{split}
%e_k &= 
E_c=\frac{ c\delta + c(1-\delta) \sigma^{1-\alpha}}
{c  - c_u }.\\
%\end{split}
\end{equation}

@@

\begin{figure}[htbp]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={PDS Coverage},
            width=\linewidth,
            height=0.6\linewidth,
            cycle multi list={
                mline\nextlist
                [3 of]mmark*\nextlist
            },
            %cycle list name=mcolor,
            xlabel={Total num. chunks stored (in billions)},
            ylabel={PDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            mark options=solid,
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.30)},
            anchor=north},
            ]
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A4] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T4] {figures/cds_coverage.txt};
            \legend{Measured ($\sigma=1\%$),Measured ($\sigma=2\%$),Measured ($\sigma=4\%$),Predicted ($\sigma=1\%$),Predicted ($\sigma=2\%$),Predicted ($\sigma=4\%$)};
            \end{axis}
    \end{tikzpicture}
  \caption{Predicted vs. actual PDS coverage as data size increases.}
  \label{fig:cds-coverage}
\end{figure}
Ratio $\sigma^{1-\alpha}$ represents the percentage of the remaining
chunks detected as duplicates in global deduplication due to PDS.
We call this PDS coverage.
Figure~\ref{fig:cds-coverage} shows predicted PDS coverage using $\sigma^{1-\alpha}$ when $\alpha$ is fixed at
0.65 and measured PDS coverage in our test dataset.
$\sigma=2\%$ represents memory usage of approximately 100MB memory per machine for the PDS.
While the predicted value remains flat, PDS  coverage actually increases as more VMs are involved.
increases. This is because actual $\alpha$ value increases with the data size.
%Then  PDS coverage PDS increases as more VMs are involved.

\subsection{ Impact Fault Isolation}

@@

The replication degree of the backup storage 
is $r$ for regular file blocks and $r=3$ is a typical setting in distributed
file systems~\cite{googlefs03,hdfs10}.
Since $\sigma$ is small (it is 2\% in our experiments),  
the impact of replication on storage increase is very small even 
when we choose a large $r_c/r$ ratio.  For example, $r_c/r=2$ or 3.

%In the VC approach, a special replication degree $r_c$ is used for PDS data where $r_c>r$. 
%The storage cost for VO with full deduplication is $c_u *r$ and for VC, it is
%$ k*r_c  + (c-E(c-c_u))*r$. Thus The storage cost ratio of $VC$ over $VO$ is 
%\[
%\sigma \frac{r_c}{r} + \frac{c-e(c-c_u)}{c_u}.
%\]
%Since $\sigma$ is small (it is 2\% in our experiments),  
%term $\sigma \frac{r_c}{r}$ is small in the above expression.  
%Thus the impact on storage increase is very small even when we choose a large $r_c/r$ ratio. 
%For example, $r_c/r=2$ or 3.

Now we  assess  the impact of losing $d$ machines 
to the the VC and VO approaches.  
A large $r_c/r$ ratio can have a positive impact on full availability of VM snapshot blocks.
We use file system blocks rather than a deduplication
data chunk as our unit of failure because the DFS keeps
filesystem blocks as its base unit of storage.
To compute the full availability of all snapshots of a VM, we derive
the probability of losing a snapshot file block of a VM by
estimating the number of file system blocks per VM in each approach.
As illustrated in Figure~\ref{fig:share},
we build a bipartite graph representing the association from unique file system blocks
to their corresponding VMs in each approach. An association edge is  drawn  from a file block  to a VM 
if this block is used by the VM.

For VC, each VM has an average number of $N_1$ non-PDS file system blocks
and  has  an average of  $N_2$ PDS file system blocks. 
Each non-PDS file system block is associated with one VM
 and  we denote that PDS file system blocks are
shared by an average of $V_c$ VMs. Let $s$ be the average number of chunks per file block, 
$V$ be the average number of VMs hosted on each machine. Then, 
\[
V *p*N_1 *s  \approx c-E_c(c-c_u) - c_u\sigma\; \mbox{ and } \; 
V *p*N_2 *s  \approx c_u \sigma V_c.
\]

For VO, each VM has an average of $N_o$ file system blocks
and let $V_o$ be the average number of VMs shared by each file system block.
\[
V *p*N_o *s  = (c- E_o(c-c_u) ) V_o.
\]

Since each file block (with default size $64MB$) contains many chunks (on average 4KB),
each file block contains the hot low-level chunks shared by many VMs, and it also contains
rare chunks which are not shared.  Since $c>> c_u$, from the above equations:
\[
\frac{N_1}{N_o} \approx  \frac{1-E_o}{(1-E_c) V_o}.
\] 
When $E_c\approx E_o$ as we demonstrate in Section~\ref{sect:evaldedup}, $N_1$ would be much smaller than $N_o$. 
When there is a failure in file blocks with replication degree $r$
and there is no failure for PDS data with more replicas,   a VM in
the VC approach has a much lower chance to lose a snapshot than VO. 
%Figure~\ref{fig:fsb-links} shows the average number of VMs sharing each file block from
%our test dataset.
%For VC, we only show
%the sharing degree of  PDS-oriented file blocs  because non-PDS blocks always belong to
%one VM.
%Figure~\ref{fig:fsb-links} shows the average number of VMs sharing each file block from our test data.
Figure~\ref{fig:vm-links} shows the average number of file system blocks for each VM in VC and in VO. 
From data, we have observed that $N_1$ is much smaller than $N_o$.  
In the evaluation discussed in Section~\ref{ref:evaldedup}, we have considered a worst case scenario that
every PDS file block is shared by all VMs in the VC approach, which leads a large $N_2$ value. 
Even with that, the availability of VC snapshots is much higher than VO and we analyze  this below.
%In fact, $N_1 +N_2 < N_o$. 
%This is likely because the PDS FSBs tightly pack data used by many VMs, 
%which decreases the overall number of FSBs required to backup a VM.
%Note that  if  the backup for multiple VMs were conducted concurrently, there could be many more
%VMs sharing each file block on average in VO.

@@

\item
When $r_c \leq d$, both non-PDS and PDS file blocks in VC can have a loss.
The full snapshot availability of  a VM in the VC approach is
\begin{equation}
\label{eq:VC2}
% (1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
% *
% (1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2}.
A(r)^{N_1} * A(r_c)^{N_2}
\end{equation}
Table~\ref{tab:fsb-availability} lists the $A(r)$ values with
%that the availability of an individual file system block
with different replication degrees. This demonstrates inequality 
% $1-\frac{ \binom{d}{r}} { \binom{p}{r} } < 1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} }$
$A(r) < A(r_c)$ because of  $r < r_c$.  
Since $N_1$ is much smaller than $N_o$ as discussed previously, 
these two conditions support that VC has a  higher availability. 
%comparing  Formula~\ref{eq:VO} and~\ref{eq:VC}. 
\end{itemize}
